\begin{enumerate}[label={[\arabic*]}]
\item Shor P W 1994 Algorithms for quantum computation: discrete logarithms and factoring \textit{IEEE Comput Soc Press}. 124

\item Grover L K 1997 Quantum Mechanics Helps in Searching for a Needle in a Haystack \textit{Phys. Rev. Lett.} \textbf{79} 325

\item Zanardi P and Rasetti M 1999 Holonomic quantum computation \textit{Phys. Lett.} A \textbf{264} 94

\item Sjöqvist E and Tong D M and Mauritz Andersson L and Hessmo B and Johansson M and Singh K 2012 Non-adiabatic holonomic quantum computation \textit{New Journal of Physics} \textbf{14} 10

\item Berry M V 1984 Quantal phase factors accompanying adiabatic changings \textit{Proceedings of the Royal Sociery of London Series A, Mathematical and Physical Sciences} \textbf{392} 1802

\item Aharonov Y and Anandan J 1987 Phase change during cyclic quantum evolution \textit{Phys. Rev. Lett.} \textbf{58} 16

\item Anandan J 1988 Non-adiabatic non-abelian geometric phase \textit{Phys. Lett.} A \textbf{133} 4

\item Wilczek F and Zee A 1984 Appearance of gauge structure in simple dynamical systems \textit{Phys. Rev. Lett} \textbf{52} 24

\item Glusker M and Hogan D M and Vass P 2005 The ternary calculating machine of Thomas Fowler \textit{IEEE Annals of the History of Computing} \textbf{27} 3

\item Brusentsov N P and Ramil Alvarez J 2011 Ternary Computers: The Setun and the Setun 70 \textbf{357} 74

\item Li B and Yu Z H and Fei S M 2013 Geometry of quantum computation with qutrits \textit{Scientific Reports} \textbf{3} 1

\item Lu H and Hu Z, Alshaykh M S and Moore A J and Wang Y and Imany P and Weiner A M and Kais S 2020 Quantum Phase Estimation with Time‐Frequency Qudits in a Single Photon \textit{Advanced quantum technologies} \textbf{3} 2

\item Luo M and Wang X 2014 Universal quantum computation with qudits \textit{Science China Physics, Mechanics \& Astronomy} \textbf{57} 9


\item Wang Y and Hu Z and Sanders B C and Kais S 2020 Qudits and High-Dimensional Quantum Computing \textit{Frontiers in Physics} \textbf{8} 589504

\item Ai M Z and Li S and He R and Xue Z Y and Cui J M, Huang Y F and Li C F and Guo G C 2021 Experimental Realization of
Nonadiabatic Holonomic Single-Qubit Quantum Gates with Two Dark Paths in
a Trapped Ion \textit{Fundamental Research}

\item Sakurai JJ and Napolitano J 2011 Modern quantum mechanics; 2nd ed. \textit{San Francisco, CA: Addison-Wesley}

\item Pavlidis A and Floratos E 2021 Quantum-Fourier-transform-based quantum arithmetic with qudits \textit{Physical Review} A \textbf{103} 2

\item Howard M and Vala J 2012 Qudit versions of the qubit gate \textit{Physiscal Review A, Atomic, Molecular, and Optical Physics} \textbf{86} 2 

\item Shkolnikov V O and Burkard G 2020 Effective Hamiltonian theoryof the geometric evolution of quantum systems \textit{Physical Review} A \textbf{101} 4

\item Born M and Fock V 1928 Beweis des Adiabatensatzes \textit{The European Physical Journal A, Hadrons and Nuclei} \textbf{51} 3

\item Duan L and Cirac J I and Zoller P 2001 Geometric Manipulation of Trapped Ion for Quantum Computations \textit{Science} \textbf{292} 5522

\item Herterich E and Sjöqvist E 2016 Single-loop multiple-pulse nonadiabatic holonomic quantum gates \textit{Physical Review} A \textbf{94} 5

\item Morris J R and Shore B W 1983 Reduction of degenerate two-level excitation to independent two-state system \textit{Physical Review A, General physics} \textbf{27} 2

\item Shampine L F and Reichelt M W 1997 The MATLAB ODE Suite \textit{SIAM journal on scientific computing} \textbf{18} 1

\item Campbell E T and Anwar H and Browne D E 2012 Magic-State Distillation in All Prime Dimensions Using Quantum Reed-Muller Codes \textit{Physical Review} X \textbf{2} 4




\end{enumerate}